A White Noise Approach to Evolutionary Ecology

### 2020-04-06 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
ALL of the rigorous proofs are finally complete

To stay within the realm of biological plausibility we require a set technical assumptions. First, we assume the initial abundance density is continuous, non-negative, integrable and has finite trait mean and variance. That is, we assume $\nu(x,0)$ is continuous in $x$, satisfies $\nu(x,0)\geq0$ for all $x\in\mathbb{R}$ and
\begin{equation}
N(0)=\int_{-\infty}^{+\infty}\nu(x,0)dx<+\infty,
\end{equation}
\begin{equation}
-\infty<\bar x(0)=\int_{-\infty}^{+\infty}xp(x,0)dx<+\infty,
\end{equation}
\begin{equation}
\sigma^2(0)=\int_{-\infty}^{+\infty}(x-\bar x(0))^2p(x,0)dx<+\infty,
\end{equation}
where $\bar x(t)$ and $\sigma^2(t)$ are respectively the mean trait and phenotypic variance at time $t\geq0$. Second, we assume selection is determined by the growth rate $m(\nu(x,t),x)$ that is differentiable with respect to both arguments and satisfies $m(y,x)\leq r$ for some $r\in\mathbb{R}$ and for all $x\in\mathbb{R}$, $y\geq0$. From here on we abbreviate $m(\nu(x,t),x)$ to $m(\nu,x)$. Third, we assume mutation is captured by diffusion with coefficient $\frac{\mu}{2}$. With these technicalities aside, the demographic dynamics can be modelled by the PDE
\begin{equation}\label{eq1}
\frac{\partial}{\partial t}\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)
\end{equation}
with the initial condition $\nu(x,0)$ described above. This PDE is semilinear due to the dependency of the growth rate $m(\nu,x)$ on the solution $\nu(x,t)$ and is referred to as a scalar reaction-diffusion equation [@lawrenceevans2010]. Interpreting the trait value $x$ as spatial location, this model of population dynamics has been intensely studied [for a review, see @Cantrell2004]. When $\mu=0$, equation (\ref{eq1}) can be seen as an analog of equation (\ref{first}) for a continuum of types. By assuming mutation acts as diffusion the effect of mutation causes $\nu(x,t)$ to flatten out over time. In fact, if the growth rate is constant across $x$, then this model of mutation will cause $\nu(x,t)$ to converge to a flat line as $t\to\infty$. Although clearly an idealized representation of biological reality, this model is sufficiently general to capture a large class of dynamics including density dependent growth and frequency dependent selection. As an example, logistic growth combined with quadratic stabilizing selection can be captured using the growth rate 
\begin{equation}
m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\nu(x,t)
\end{equation}
where $a\geq0$ is the strength of stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c\geq0$ captures the sensitivity of fitness to intraspecific competition and $r\in\mathbb{R}$ is the intrinsic growth rate in the absence of abiotic selection. This exemplary fitness function has a few convenient properties that we breifly take note of. First, the effect of competition induces a carrying capacity on the population, leading to a finite equilibrium abundance over bounded subsets of $\mathbb{R}$. Second, abiotic selection prevents the abundance density from diffusing too far from the abiotic optimum, leading to a finite equilibrium abundance across all of $\mathbb{R}$. A stochastic generalization of this model, closely related to our model of diffuse coevolution presented later, has been studied by Dawson [-@Dawson1978].

@@

In SM \S\ref{SDE_DERIV}, we extend these results to include the effects of demographic stochasticity. Similar to the approach taken by Champagnat, Ferriére and Méléard [-@Champagnat2006], we begin with a BBM that models populations as discrete sets of reproducing individuals whose vital rates depend on their trait value as well as the state of the entire population. Taking a large population size limit and keeping our assumption of single dimensional traits, we employ a pair of classical results that show, under the appropriate rescaling in time, phenotypic space and population density, a sequence of rescaled BBM converges to a limiting process that can be characterized by a SPDE [@Mlard1993; @zeng1998absolute]. The limiting processes of rescaled BBM have been referred to as measure-valued Markov processes [@dawson1993measure] or superprocesses [@alisonetheridge2000]. Under the simplifying assumptions inherited from our treatment of deterministic dynamics and an additional assumption on demographic stochasticity, we obtain as a special case a relatively simple expression for an SPDE that generalizes equation (\ref{eq1}). The simplicity of our special case allows us to use properties of space-time white noise processes to derive a set of SDE that generalize equations (\ref{N_det}), (\ref{detmean}) and (\ref{vardyn_det}) to include the effects of demographic stochasticity. Classical expressions for the effects of random genetic drift on the evolution of mean traits are obtained as a special case.

In the following section we provide the necessary mathematical tools needed to derive SDE from SPDE. Since our aim is to present this material to as wide of audience as possible, our approach deviates from standard definitions to remove the need for a detailed technical treatment. In addition to the notions of Reimann integration and partial differentiation already employed, the reader will only need some elementary probability and an intuitive understanding of SDE, including Brownian motion. Because space-time white noise, denoted by $\dot W(x,t)$, appears in the SPDE characterizing diffusion limits of BBM, we begin by defining $\dot W(x,t)$ and illustrating its relevant properties including a set of heuristics for performing calculations. Treating only the simplist of cases, we then provide a brief review of BBM, their diffusion limits and the SPDE that characterize them. For those not interested in the white noise calculus or superprocesses and would rather jump straight into more biologically relevant results, we recommend skipping to \S\ref{particular}.

@@

Throughout this supplement, we set use dot notation for time derivatives so that $\dot f(x,t)=\frac{\partial }{\partial t}f(x,t)$ and set $\Delta=\frac{\partial^2}{\partial x^2}$, except in \S\ref{caC} where $\Delta$ represents a random variable.

@@

Recall that $m(\nu,x)$ is shorthand for $m(\nu(x,t),x)$. That is, $m:[0,+\infty)\times\mathbb{R}\to\mathbb{R}$. Following our assumptions of the main text, we have that $m(y,x)$ is differentiable with respect to both $x$ and $y$ and there exists $r\in\mathbb{R}$ such that $m(y,x)\leq r$ across all $y\geq0$ and $x\in\mathbb{R}$. Hence, we can apply Proposition 1.21 of Cantrell and Cosner [-@Cantrell2004] to show that $\nu(x,t)$ remains differentiable with respect to $t$ and twice differentiable with respect to $x$ for all $t\geq0$.

As in the main text, we also assume the initial condition $\nu(x,0)$ satisfies
\begin{equation}
N(0)=\int_{\mathbb{R}}\nu(x,0)dx<+\infty,
\end{equation}
\begin{equation}
-\infty<\bar x(0)=\int_{\mathbb{R}}xp(x,0)dx<+\infty,
\end{equation}
\begin{equation}
\sigma^2(0)=\int_{\mathbb{R}}(x-\bar x(0))^2p(x,0)dx<+\infty,
\end{equation}
where $p(x,0)=\nu(x,0)/N(0)$, and we consider the PDE
\begin{equation}\label{PDE_SM}
\dot\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t).
\end{equation}

@@

@@

@@

\begin{multline}
\dot{\sigma}^2(t)=\frac{d}{dt}\int_\mathbb{R}(x-\bar x(t))^2p(x,t)dx=\int_\mathbb{R}2(x-\bar x(t))\dot{\bar x}(t)+(x-\bar x(t))^2\dot p(x,t)dx\\=\int_\mathbb{R}(x-\bar x(t))^2\left((m(\nu,x)-\bar m(t))p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)\right)dx\\=\int_\mathbb{R}\left((x-\bar x(t))^2-\sigma^2(t)+\sigma^2(t)\right)(m(\nu,x)-\bar m(t))p(x,t)+(x-\bar x(t))^2\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)dx\\=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx.
\end{multline}
Applying integration by parts twice yields
\begin{equation}
\int_{-\infty}^{+\infty}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx=2.
\end{equation}

@@

\begin{multline}
\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)=\sigma^2\left(\frac{\partial}{\partial\bar x}\int_\mathbb{R}m(\nu,x)p(x,t)dx-\int_\mathbb{R}p(x,t)\frac{\partial}{\bar x}m(\nu,x)dx\right)\\=\sigma^2\int_\mathbb{R}m(\nu,x)\frac{\partial}{\partial\bar x}p(x,t)dx=\sigma^2\int_\mathbb{R}\frac{x-\bar x(t)}{\sigma^2}m(\nu,x)p(x,t)dx\\=\int_\mathbb{R}(x-\bar x)(m(\nu,x)-\bar m)p(x,t)dx=\mathrm{Cov}_t(m,x),
\end{multline}

@@

the density process $\nu(x,t)$ defined by SPDE (\ref{SPDE}) of the main text satisfies $\int_\mathbb{R}(|x|+x^2)\nu(x,t)dx<\infty$. From SM \S\ref{finite}, under the assumption $m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(x,t)dx$, we can derive the differential equations

\begin{subequations}
\begin{equation}
\dot{\bar x}=aG(\theta-\bar x)
\end{equation}
\begin{equation}
\dot G=\mu-aG^2
\end{equation}
\begin{equation}
\dot{N}=\left(r-\frac{a}{2}\left((\theta-\bar x)^2+G+\eta\right)-cN\right)N.
\end{equation}
\end{subequations}

Ignoring the trivial case of $N=0$, the equilibrium

\begin{subequations}
\begin{equation}
\hat{\bar x}=\theta,
\end{equation}
\begin{equation}
\hat G=\sqrt\frac{\mu}{a},
\end{equation}
\begin{equation}
\hat N=\frac{1}{c}\left(r-\frac{1}{2}(\eta a+\sqrt{\mu a})\right),
\end{equation}
\end{subequations}

is unique and globally stable for $a,c,\mu>0$. Setting $2r>\eta a+\sqrt{\mu a}$ ensures a positive equilibrium abundance and setting $c<r-(\eta a+\sqrt{\mu a})/2$ ensures $\hat N>1$, which is important for numerical simulations when $N$ is an integer. We use these results to help guide our choice of parameter values for simulations of the branching random walk. In the following section we provide a detailed description of the branching random walk and how we have chosen to rescale it. We then use the rescaled branching random walk to investigate finiteness of moments and normality.

>>>>

To stay within the realm of biological plausibility we require a set technical assumptions. First, we assume the initial abundance density is continuous, non-negative, integrable and has finite trait mean and variance. That is, we assume $\nu(x,0)$ is continuous in $x$, satisfies $\nu(x,0)\geq0$ for all $x\in\mathbb{R}$ and
N(0)=\int_{-\infty}^{+\infty}\nu(x,0)dx<+\infty,
-\infty<\bar x(0)=\int_{-\infty}^{+\infty}xp(x,0)dx<+\infty,
\sigma^2(0)=\int_{-\infty}^{+\infty}(x-\bar x(0))^2p(x,0)dx<+\infty,
where $\bar x(t)$ and $\sigma^2(t)$ are respectively the mean trait and phenotypic variance at time $t\geq0$. Second, we assume selection is determined by the growth rate $m(h,x)$ that is differentiable with respect to both arguments and is bounded above by some $r\in\mathbb{R}$ for all $x\in\mathbb{R}$ and $h\geq0$. We denote by $K$ an operator that takes functions such as $\nu(x,t)$ as its argument and returns other functions that may still depend on the spatial variable $x$ and the temporal variable $t$. For our application we will be focused on operators $K$ that can be written as $K(\nu)(x,t)=\int_{-\infty}^{+\infty}\kappa(x,y)\nu(y,t)dy$ for some non-negative and bounded bivariate function $\kappa$. We set $h=K(\nu)(x,t)$ to account for nonlocal effects of the abundance density $\nu(x,t)$ on the fitness of individuals with trait value $x$ [@Volpert2014]. An example of a nonlocal effect is the competition for a common resource among all individuals regardless of their trait value so that individuals with differing trait values effect each others fitness. The term nonlocal comes from interpreting the trait value as spatial location. From here on we abbreviate $m(h,x)=m(K(\nu)(x,t),x)$ to just $m(\nu,x)$. Third, we assume mutation is captured by diffusion with coefficient $\frac{\mu}{2}$. Now that these technicalities have been treated we can model the demographic dynamics of a population and the dynamics of a quantitative character simultaneously by the PDE
\begin{equation}\label{eq1}
\frac{\partial}{\partial t}\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)
with the initial condition $\nu(x,0)$ described above. This PDE is semilinear due to the dependency of the growth rate $m(\nu,x)$ on the solution $\nu(x,t)$ and is referred to as a scalar reaction-diffusion equation [@lawrenceevans2010]. Interpreting the trait value $x$ as spatial location and ignoring nonlocal effects, which can be satisfied by setting $K(\nu)=\nu$, this model of spatially distributed population dynamics has been intensely studied [for a review, see @Cantrell2004]. When $\mu=0$, equation (\ref{eq1}) can be seen as an analog of equation (\ref{first}) for a continuum of types. By assuming mutation acts as diffusion, the effect of mutation causes $\nu(x,t)$ to flatten out over time. In fact, if the growth rate is constant across $x$, then this model of mutation will cause $\nu(x,t)$ to converge to a flat line as $t\to\infty$. Although clearly an idealized representation of biological reality, this model is sufficiently general to capture a large class of dynamics including density dependent growth and frequency dependent selection. As an example, logistic growth combined with quadratic stabilizing selection can be captured using the growth rate 
m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_{-\infty}^{+\infty}\nu(y,t)dy
where $r\in\mathbb{R}$ is the intrinsic growth rate in the absence of abiotic selection, $a\geq0$ is the strength of stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c\geq0$ captures the sensitivity of fitness to intraspecific competition and we have set $\kappa(x,y)=1$. This exemplary fitness function has a few convenient properties that we breifly take note of. First, the effect of competition induces a local carrying capacity on the population, leading to a finite equilibrium abundance over bounded subsets of $\mathbb{R}$. Second, abiotic selection prevents the abundance density from diffusing too far from the abiotic optimum, leading to a finite equilibrial abundance across all of $\mathbb{R}$, $\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})$, finite equilibrial mean trait $\lim_{t\to\infty}\bar x(t)=\theta$ and finite equilibrial phenotypic variance $\lim_{t\to\infty}\sigma^2(t)=\sqrt{\mu/a}$, so long as $|\bar x(0)|<+\infty$, $\sigma^2(0),N(0)>0$ and $r\geq\sqrt{a\mu}/2$. We prove these statements in SM \S\ref{equilib}. By replacing $\kappa(x,y)=1$ with a particular Gaussian function, we will find in \S\ref{coev} and SM \S\ref{diffuse} this fitness function becomes equivalent to a fitness function derived from niche theory that we use to model coevolution driven by resource competition.

@@

In SM \S\ref{SDE_DERIV}, we extend these results to include the effects of demographic stochasticity. Similar to the approach taken by Champagnat, Ferriére and Méléard [-@Champagnat2006], we begin with a BBM that models populations as discrete sets of reproducing individuals whose vital rates depend on their trait value as well as the state of the entire population. Taking a large population size limit and keeping our assumption of single dimensional traits, we employ a pair of classical results that show, under the appropriate rescaling in time, phenotypic space and population density, a sequence of rescaled BBM converges to a limiting process that can be characterized by a SPDE [@Mlard1993; @zeng1998absolute]. The limiting processes of rescaled BBM have been referred to as measure-valued Markov processes [@dawson1993measure] or superprocesses [@alisonetheridge2000]. Under the simplifying assumptions inherited from our treatment of deterministic dynamics and the additional assumption that the magnitude of demographic stochasticity is independent of trait values, we obtain as a special case a relatively simple expression for an SPDE that generalizes equation (\ref{eq1}). The simplicity of our special case allows us to use properties of space-time white noise processes to derive a set of SDE that generalize equations (\ref{N_det}), (\ref{detmean}) and (\ref{vardyn_det}) to include the effects of demographic stochasticity. Classical expressions for the effects of random genetic drift on the evolution of mean traits are obtained as a special case.

In the following section we provide the necessary mathematical tools needed to derive SDE from SPDE. Since our aim is to present this material to as wide of audience as possible, our approach deviates from standard definitions to remove the need for a detailed technical treatment. In addition to the notions of Reimann integration and partial differentiation already employed, the reader will only need some elementary probability and an intuitive understanding of SDE, including Brownian motion. Because space-time white noise, denoted by $\dot W(x,t)$, appears in the SPDE characterizing diffusion limits of BBM, we begin by defining $\dot W(x,t)$ and illustrating its relevant properties including a set of heuristics for performing calculations. Treating only the simplist of cases, we then provide a brief review of BBM, their diffusion limits and the SPDE that characterize them. For those not interested in the white noise calculus or superprocesses and would rather jump straight into more biologically relevant results, we recommend skipping to \S\ref{equations}.

@@

Throughout this supplement, we set use dot notation for time derivatives so that $\dot f(x,t)=\frac{\partial }{\partial t}f(x,t)$ and set $\Delta=\sum_{i=1}^d\frac{\partial^2}{\partial x_i^2}$ the Laplace operator on $\mathbb{R}^d$, except in \S\ref{caC} where $\Delta$ represents a random variable.

@@

Recall that $m(\nu,x)$ is shorthand for $m(\nu(x,t),x)$. That is, $m:[0,+\infty)\times\mathbb{R}\to\mathbb{R}$. Following our assumptions of the main text, we have that $m(y,x)$ is differentiable with respect to both $x$ and $y$ and there exists $r\in\mathbb{R}$ such that $m(y,x)\leq r$ across all $y\geq0$ and $x\in\mathbb{R}$. As in the main text, we also assume the initial condition $\nu(x,0)$ satisfies
\begin{equation}
N(0)=\int_{\mathbb{R}}\nu(x,0)dx<+\infty,
\end{equation}
\begin{equation}
-\infty<\bar x(0)=\int_{\mathbb{R}}xp(x,0)dx<+\infty,
\end{equation}
\begin{equation}
\sigma^2(0)=\int_{\mathbb{R}}(x-\bar x(0))^2p(x,0)dx<+\infty,
\end{equation}
where $p(x,0)=\nu(x,0)/N(0)$, and we consider the PDE
\begin{equation}\label{PDE_SM}
\dot\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t).
\end{equation}

@@

## Regularity of deterministic solution \label{regular}

Here we show that solutions to the PDE described in the main text are continuous in time and space, twice continuously differentiable with respect to space and differentiable with respect to time. Our approach is an adaptation of that taken by Volpert [-@Volpert2014].

Recall that the PDE we consider is of the form
\begin{equation}\label{PDE_supp}
\dot\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t),
\end{equation}
where $m(\nu,x)$ bounded above by $r\in\mathbb{R}$ is shorthand for $m(K(\nu)(x,t),x)$ and $K(\nu)(x,t)=\int_\mathbb{R}\kappa(x,y)\nu(y,t)dy$ for some non-negative, bounded $\kappa$. Denoting $L^1(\mathbb{R})$ the Lebesgue space of integrable functions on $\mathbb{R}$ (ie, the set of functions $f:\mathbb{R}\to\mathbb{R}$ that satisfy $\|f\|_1=\int_\mathbb{R}|f(x)|dx<+\infty$ under the equivalence relation $f=g$ iff $\|f-g\|_1=0$), we assume $\nu(\cdot,0)\in L^1(\mathbb{R})$. Under these assumptions, we have already proved in SM \S\ref{finite} that $\nu(\cdot,t)\in L^1(\mathbb{R})$ for all $t\geq0$. Before proving the regularity of $\nu(x,t)$, we review the approach taken by Volpert [-@Volpert2014].

__Volperts approach__

In chapter 9 \S2.2 of Volpert [-@Volpert2014] complimentary conditions are assumed where $\kappa(x,y)=\varphi(x-y)$ for some integrable function $\varphi$ and initial condition $\nu(\cdot,0)\in L^\infty(\mathbb{R})$, where $L^\infty(\mathbb{R})$ is the Lebesgue space of bounded functions on $\mathbb{R}$ (ie, the set of functions $f:\mathbb{R}\to\mathbb{R}$ satisfying $\|f\|_\infty=\sup_{x\in\mathbb{R}}|f(x)|<+\infty$ under the equivalence relation $f=g$ iff $\|f-g\|_\infty=0$). Instead of the growth term $m(\nu,x)\nu(x,t)$ of PDE (\ref{PDE_supp}), Volpert worked with $f(\nu(x,t),K(\nu)(x,t))$ for some function $f:\mathbb{R}^2\to\mathbb{R}$ differentiable with respect to both arguments and satisfying, for some $q\geq0$,
\begin{equation}
f(x,y)\leq q(1+|x|+|y|), \ \forall \ x,y\in\mathbb{R}.
\end{equation}
Under these conditions, Volpert used semigroup theory to prove the solution $\nu(x,t)$ takes values in $L^\infty(\mathbb{R})$ for all $t\geq0$ and is differentiable with respect to time and twice continuously differentiable with respect to space. In particular, defining $\nu(t)$ as the function $x\to\nu(x,t)$ and keeping $\Phi(x,t)$ as defined in SM \S\ref{finite}, Volpert states the semigroup $\{S(t)\}_{t\geq0}$ given by 
\begin{equation}
S(t)\nu(0)=\int_0^t\frac{\mu}{2}\Delta\nu(x,s)ds=\int_\mathbb{R}\Phi(x-y,t)\nu(y,0)dy, \ \ \forall \ \ t\geq0, \ \nu(0)\in L^\infty(\mathbb{R})
\end{equation}
satisfies $\|S(t)\nu(0)\|_\infty\leq\|\nu(0)\|_\infty$ for all $t\geq0$ and $\nu(0)\in L^\infty(\mathbb{R})$. Then, by setting
\begin{equation}
F(v)=f\left(v(x),\int_\mathbb{R}\varphi(x-y)v(y)dy\right), \ \ \forall \ \ v\in L^\infty(\mathbb{R}),
\end{equation}
Volpert notes that $F$ is locally Lipschitz continuous on $L^\infty(\mathbb{R})$ and in particular $\|F(v)\|_\infty\leq q\|v\|_\infty$. Hence, under Volperts conditions, solutions to PDE (\ref{PDE_supp}) can be written as
\begin{equation}\label{volp_sol}
\nu(t)=S(t)\nu(0)+\int_0^tS(t-s)F(\nu(s))ds\in L^\infty(\mathbb{R}) \ \ \forall \ \ t\geq0, \ \nu(0)\in L^\infty(\mathbb{R}).
\end{equation}
Thanks to properties of the semigroup $\{S(t)\}_{t\geq0}$, solutions $\nu(t)$ given by (\ref{volp_sol}) satisfy the claimed regularity conditions.

__Adapting Volperts approach to our context__

To adapt Volperts approach to our situation, note that $\|S(t)\nu(0)\|_1=\|\nu(0)\|_1$ for all $\nu(0)\in L^1(\mathbb{R})$ and, setting
\begin{equation}
F(v)=m\left(\int_\mathbb{R}\kappa(x,y)v(y)dy,x\right)v(x), \ \ \forall \ \ v\in L^1(\mathbb{R}),
\end{equation}
we have $\|F(v)\|_1\leq r\|v\|_1$ for all $v\in L^1(\mathbb{R})$. Hence, using the same arguments employed by Volpert [-@Volpert2014], we have that solutions to PDE (\ref{PDE_supp}) under our assumptions can be written as
\begin{equation}
\nu(t)=S(t)\nu(0)+\int_0^tS(t-s)F(\nu(s))ds\in L^1(\mathbb{R}) \ \ \forall \ \ t\geq0, \ \nu(0)\in L^1(\mathbb{R})
\end{equation}
which therefore satisfy the claimed regularity conditions.

@@

## Equilibrium moments for a deterministic population experiencing logistic growth and stabilizing selection \label{equilib}

Here we set out to show, given appropriate initial conditions, $N_\infty=\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})$, $\bar x_\infty=\lim_{t\to\infty}\bar x(t)=\theta$ and $\sigma^2_\infty=\lim_{t\to\infty}\sigma^2(t)=\sqrt{\tfrac{\mu}{a}}$ when
\begin{equation}\label{m_log_stab}
m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(y,t)dy
\end{equation}
for $\theta\in\mathbb{R}$, $\mu\geq0$, $a,c>0$ and $r\geq\tfrac{1}{2}\sqrt{\mu a}$. Set $\nu_\infty(x)=\lim_{t\to\infty}\nu(x,t)$ and $p_\infty(x)=\nu_\infty(x)/N_\infty$, when these quantities exist. The mean fitness can be calculated as
\begin{equation}
\bar m(t)=r-\frac{a}{2}((\theta-\bar x(t))^2-\sigma^2(t))-cN(t), \ \ \forall \ \ t\geq0.
\end{equation}
Solving for $N_\infty$ amounts to setting $\dot N(t)=0$ and solving for $N(t)$. This reduces to solving $\bar m(t)=0$ for $N(t)$, which returns
\begin{equation}
N_\infty=\frac{1}{c}\left(r-\frac{a}{2}\left((\theta-\bar x_\infty)^2-\sigma^2_\infty\right)\right).
\end{equation}
To find $\nu_\infty(x)$ we solve $\tfrac{\partial}{\partial t}\nu(x,t)=0$ for $\nu(x,t)$. This implies the following ordinary differential equation
\begin{equation}
\frac{d^2}{dx^2}\nu_\infty(x)=\left(\frac{2c}{\mu}N_\infty+\frac{a}{\mu}(\theta-x)^2-\frac{2r}{\mu}\right)\nu_\infty(x)
\end{equation}
which has the solution
\begin{equation}
\nu_\infty(x)=\frac{N_\infty}{\sqrt{2\pi}}\left(\frac{a}{\mu}\right)^{\frac{1}{4}}\exp\left(-\frac{(\theta-x)^2}{2\sqrt{\mu/a}}\right).
\end{equation}
From this expression we infer $\bar x_\infty=\theta$ and $\sigma^2_\infty=\sqrt{\mu/a}$. Hence $N_\infty=\frac{1}{c}\left(r-\frac{1}{2}\sqrt{a\mu}\right)$. This holds for any initial $\nu(\cdot,0)\in L^1(\mathbb{R})$ such that $\bar x(0)\in\mathbb{R}, \ \sigma^2(0),N(0)>0$.

@@

\begin{multline}
\frac{d}{dt}\sigma^2(t)=\frac{d}{dt}\int_\mathbb{R}(x-\bar x(t))^2p(x,t)dx=\int_\mathbb{R}2(x-\bar x(t))\frac{d}{dt}{\bar x}(t)+(x-\bar x(t))^2\frac{\partial}{\partial t} p(x,t)dx\\=\int_\mathbb{R}(x-\bar x(t))^2\left((m(\nu,x)-\bar m(t))p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)\right)dx\\=\int_\mathbb{R}\left((x-\bar x(t))^2-\sigma^2(t)+\sigma^2(t)\right)(m(\nu,x)-\bar m(t))p(x,t)+(x-\bar x(t))^2\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t)dx\\=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)+\frac{\mu}{2}\int_\mathbb{R}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx.
\end{multline}
Applying integration by parts twice yields
\begin{equation}
\int_{-\infty}^{+\infty}(x-\bar x(t))^2\frac{\partial^2}{\partial x^2}p(x,t)dx=2.
\end{equation}

@@

\begin{multline}
\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right)=\sigma^2\left(\frac{\partial}{\partial\bar x}\int_\mathbb{R}m(\nu,x)p(x,t)dx-\int_\mathbb{R}p(x,t)\frac{\partial}{\partial \bar x}m(\nu,x)dx\right)\\=\sigma^2\int_\mathbb{R}m(\nu,x)\frac{\partial}{\partial\bar x}p(x,t)dx=\sigma^2\int_\mathbb{R}\frac{x-\bar x(t)}{\sigma^2}m(\nu,x)p(x,t)dx\\=\int_\mathbb{R}(x-\bar x)(m(\nu,x)-\bar m)p(x,t)dx=\mathrm{Cov}_t(m,x),
\end{multline}

@@

the density process $\nu(x,t)$ defined by SPDE (\ref{SPDE}) of the main text satisfies $\int_\mathbb{R}(|x|+x^2)\nu(x,t)dx<\infty$.

and setting $c<r-(\eta a+\sqrt{\mu a})/2$ ensures $\hat N>1$, which is important for numerical simulations when $N$ is an integer. We use these results to help guide our choice of parameter values for simulations of the branching random walk. In the following section we provide a detailed description of the branching random walk and how we have chosen to rescale it. We then use the rescaled branching random walk to investigate finiteness of moments and normality.
